Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-04-18 manuscript/manuscript.tex
Improved density-based discretization algorithm description in manuscript

High variations in the density function of an arbitrary tesseroid 
introduces a new type of error into the numerical approximation: only a 
few point masses would not reproduce the effect of this kind of 
variations.
The modified version of the adaptive discretization algorithm may help 
in reducing this kind of error, but as it only takes into account the 
dimensions of the tesseroid and the distance to the computation point, 
it's not well suited to fully perform this task.

In order to overcome this, we developed a complementary discretization 
algorithm that takes into account the variation of the density inside 
the tesseroid. Basically, it subdivides any tesseroid on the radial 
direction on the depth at which the \emph{maximum density variation} 
takes place.

@@

Because the algorithm will be iteratively applied for every 
subdividing tesseroid, we'd like to note that this normalised density 
function will be the same for every subdividing tesseroid.
In case that the density function is constant, both the maximum and 
minimum densities are equal and the normalised function could not 
be defined, so the density-based discretization algorithm won't be 
applied.

@@

Finally, if the product of the maximum of the absolute difference and a 
tesseroid size ratio is greater than a predefined delta ratio $\delta$ 
we will divide the tesseroid on the radius at which this maximum takes 
place ($R_\text{max}$):

@@

\noindent obtaining a pair of smaller tesseroids with the same 
longitude and latitude boundaries.

The $L_r^\text{sm}$ is the radial dimension of the tesseroid that is 
currently analysed. On the first iteration, this tesseroid is the 
\emph{original} one, so $L_r^\text{sm} = L_r$, but for future 
iterations this radial dimension will be computed as the difference of 
the top and bottom radii of the subdividing tesseroid.

This algorithm will be subsequently applied to every subdividing 
tesseroid.
We will start from equation \ref{eq:density-reference-line} defining a 
new linear function $\rho_l(r')$, an absolute difference with equation 
\ref{eq:density-abs-diff} and evaluate the discretization inequation 
\ref{eq:delta-density}, taking into account that now $R_1$ and $R_2$ 
are the radial boundaries of this smaller tesseroid.

The tesseroid size ratio included in the inequation 
\ref{eq:delta-density} is intended to reduce the number of probable 
unneeded discretizations and focus on discretizations that will 
significantly reduce the numerical error.
A big subdividing tesseroid with small maximum absolute difference would 
introduce more numerical error than a small subdividing tesseroid with 
higher maximum absolute difference.

If on the contrary the inequation \ref{eq:delta-density} is not held, 
then the tesseroid will not be subdivided.
The same happens if the tesseroid radial dimension is bellow the 
threshold of 1mm. On the first case due to the fact that no more 
discretizations are needed, and on the second one because smaller 
tesseroids will not introduce meaningful increases in the accuracy of 
the computation.
This two conditions guarantee that the algorithm is finite.
Once they are met, every subdividing tesseroid is subjected to the 
modified adaptive discretization algorithm and the GLQ in order to 
compute the gravitational field generated by it.

On Figure \ref{fig:density-discretization-algorithm}(a) we propose an 
arbitrary normalised density function as an example of how this 
algorithm works.
And we plotted its corresponding linear function $\rho_l(r')$ and a 
dashed line that represents the maximum absolute difference $\Delta 
\rho (r')$.
We will assume that the discretization inequation 
\ref{eq:delta-density} holds in this case, so we must divide the 
tesseroid at the height at which the maximum difference occurs.
The dots on Figure \ref{fig:density-discretization-algorithm} 
represent the discretization heights, i.e. the radial boundaries of the 
subdividing tesseroids produced by the algorithm.

On Figure \ref{fig:density-discretization-algorithm}(b) we apply the 
algorithm on the shallower subdividing tesseroid.
In the same way, we propose the linear function and compute the maximum 
absolute difference $\Delta \rho (r')$.
In this case, the tesseroid size-ratio $L_r^\text{sm}/L_r$ is lower 
than one and will be taken into account in the discretization 
inequation, which will be assumed to hold also in this case, leading to 
the division of this tesseroid.

>>>>

Variations in the density function of an arbitrary tesseroid introduces 
a new type of error into the numerical approximation: only a few point 
masses would not reproduce the effect of the density variations.
The modified version of the adaptive discretization algorithm may help 
in reducing this kind of error, but as it only takes into account the 
dimensions of the tesseroid and the distance to the computation point, 
it's not well suited to fully perform this task.

In order to overcome this, we developed a complementary discretization 
algorithm that takes into account the variation of the density inside 
the tesseroid. Basically, it subdivides any tesseroid on the radial 
direction on the depth at which the \emph{maximum density variation} 
takes place, only if a certain discretization inequation holds.

@@

Because the algorithm will be iteratively applied for every 
subdividing tesseroid, we'd like to note that this normalised density 
function will be the same for every subdividing tesseroid.

In case that the density function is constant, both the maximum and 
minimum densities are equal and the normalised function could not 
be defined, so the density-based discretization algorithm won't be 
applied.

@@

Finally, if the following inequation holds, the tesseroid will be divided on 
the radius $R_\text{max}$ at which the maximum absolute difference takes 
place:

@@

\noindent where the delta ratio $\delta$ is a predefined constant value 
and $L_r^\text{sm}/L_r$ is a tesseroid size ratio, where 
$L_r^\text{sm}$ and $L_r$ are the radial dimensions of the tesseroid 
that is currently analysed and of the \emph{original} one, 
respectively.
On the first iteration, the former tesseroid corresponds to the 
\emph{original} one, so $L_r^\text{sm} = L_r$, but for future 
iterations this radial dimension will be computed as the difference of 
the top and bottom radii of the subdividing tesseroid.
The tesseroid size ratio is intended to reduce the number of probable 
unneeded discretizations and focus on discretizations that will 
significantly reduce the numerical error.
A big subdividing tesseroid with small maximum absolute difference would 
introduce more numerical error than a small subdividing tesseroid with 
higher maximum absolute difference.

Once the \emph{original} tesseroid is divided, we proceed to 
subsequently apply the same algorithm on every subdividing tesseroid.
We will start from equation \ref{eq:density-reference-line} defining a 
new linear function $\rho_l(r')$, an absolute difference with equation 
\ref{eq:density-abs-diff} and evaluate the discretization inequation 
\ref{eq:delta-density}, taking into account that now $R_1$ and $R_2$ 
are the radial boundaries of this smaller tesseroid.
If the inequation holds, the current tesseroid is discretized and the 
algorithm is then applied to the resulting subdividing tesseroids.

If on the contrary the inequation \ref{eq:delta-density} is not held, 
then the tesseroid will not be subdivided.
The same happens if the tesseroid radial dimension is bellow the 
threshold of 1mm. On the first case due to the fact that no more 
discretizations are needed, and on the second one because smaller 
tesseroids will not introduce meaningful increases in the accuracy of 
the computation.
This two conditions guarantee that the algorithm is finite.
Once any of them is met and no more discretizations are carried out, 
every subdividing tesseroid is subjected to the modified adaptive 
discretization algorithm and then a second order GLQ must be applied in 
order to compute the gravitational field generated by it.

On Figure \ref{fig:density-discretization-algorithm}(a) we propose an 
arbitrary normalised density function as an example of how this 
algorithm works.
We plotted its corresponding linear function $\rho_l(r')$ and a 
dashed line that represents the maximum absolute difference.
We will assume that the discretization inequation 
\ref{eq:delta-density} holds in this case, so we must divide the 
tesseroid at the depth at which the maximum difference occurs.
The dots on Figure \ref{fig:density-discretization-algorithm} 
represent the discretization points, i.e. the radial boundaries of the 
subdividing tesseroids produced by the algorithm.

On Figure \ref{fig:density-discretization-algorithm}(b) we apply the 
algorithm on the shallower subdividing tesseroid.
In the same way, we plot the linear function and compute the maximum 
absolute difference $\Delta \rho (r')$.
In this case, the tesseroid size-ratio $L_r^\text{sm}/L_r$ is lower 
than one and will be taken into account in the discretization 
inequation, which will be assumed to hold also in this case, leading to 
the division of this tesseroid.
